---
title: "Project 2 - Team 12"
output: pdf_document
date: "2025-03-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project 2 - Team 12

# Needed packages

```{r message=FALSE, include=FALSE}
library(dplyr)
library(zoo)
library(ggplot2)
library(gridExtra)
library(frenchdata)
library(lubridate)
library(tidyr)
library(stringr)
library(zoo)
library(gt)
library(openxlsx)
library(quantmod)
library(xts)
library(TTR)
library(gridExtra)


```

# Exercise 6: Estimate the rolling Equity Beta 

To determine the Equity Beta, we use the Russel3000 Index, which includes Alnylam Pharmaceuticals, as market portfolio estimate. We use the following rolling formula to determine the Equity Beta (coming from the CAPM), which measures the sensitivity of a stock's returns relative to the market portfolio:


$$
\beta_t = \frac{\text{Cov}(R_{e,t}, R_{m,t})}{\text{Var}(R_{m,t})}
$$

where:

  - \( \beta_t \) is the equity beta at time t,
  - $\text{Cov}(R_e, R_m)$ is the covariance between the return on the asset $R_e$ and the return of the market portfolio $R_m$,
  - $\text{Var}(R_m)$ is the variance of the market return $R_m$.

### Download and clean the data:

```{r}
ALNY_PX <- read.xlsx("Daily_Data_Case_1.xlsx", sheet = 2) 
ALNY_PX$Date <- as.Date(ALNY_PX$Date,origin = "1899-12-30")

RUS_3_PX <- read.xlsx("Daily_Data_Case_1.xlsx", sheet = 1) 
RUS_3_PX$Date <- as.Date(RUS_3_PX$Date,origin = "1899-12-30")

# Observe NA values

ALNY_NA <- sum(is.na(ALNY_PX$Last.Price))
RUS_3_NA <- sum(is.na(RUS_3_PX$Last.Price))

# One NA in ALNY_PX --> Lin. Interpolate

na <- which(is.na(ALNY_PX$Last.Price))

ALNY_PX$Last.Price[na] <- (ALNY_PX$Last.Price[na-1] + ALNY_PX$Last.Price[na+1]) * 0.5


# Combine the data frames to match the dates

Prices <- inner_join(ALNY_PX , RUS_3_PX , by = "Date")
colnames(Prices) <- c("Date", "ALNY", "RUS3")

# Create xts for quantmod
Prices_xts <- xts(Prices[,-1], order.by = as.Date(Prices$Date))

```


### Equity Beta with daily returns

```{r}

ALNY_ret_d <- periodReturn(Prices_xts$ALNY, period = "daily", type = "arithmetic")
ALNY_ret_d <- data.frame(Date = index(ALNY_ret_d), coredata(ALNY_ret_d))

RUS3_ret_d <- periodReturn(Prices_xts$RUS3, period = "daily", type = "arithmetic")
RUS3_ret_d <- data.frame(Date = index(RUS3_ret_d), coredata(RUS3_ret_d))

Returns_daily <- inner_join(RUS3_ret_d  , ALNY_ret_d  , by = "Date")[-1,]
colnames(Returns_daily) <- c("Date", "ALNY", "RUS3")

start_date <- which(Returns_daily$Date == as.Date("2019-01-02")) - 200
Returns_daily <- Returns_daily[-(1:(start_date)),]

# Market Return Variance

Returns_daily$RUS3_var <- rollapply(Returns_daily$RUS3, width = 200, FUN = var, align = "right", fill = NA)

Returns_daily$rolling_cov <- runCov(
  x = Returns_daily$ALNY, 
  y = Returns_daily$RUS3, 
  n = 200
)

Returns_daily <- Returns_daily %>%
  mutate(
    E_Beta_daily = rolling_cov / RUS3_var
  )

# Remove na rows
Returns_daily <- na.omit(Returns_daily)

ggplot(Returns_daily, aes(x = Date, y = E_Beta_daily)) +
  geom_line(linewidth = 1.2,color = "blue") + 
  labs(title = "Daily Rolling Equity Beta of Alnylam Pharmaceuticals  [01/2019-12/2024]", 
       x = "Date", 
       y = "Equity Beta (200 days window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Equity Beta for weekly returns

```{r}

ALNY_ret_w <- periodReturn(Prices_xts$ALNY, period = "weekly", type = "arithmetic")
ALNY_ret_w <- data.frame(Date = index(ALNY_ret_w), coredata(ALNY_ret_w))

RUS3_ret_w <- periodReturn(Prices_xts$RUS3, period = "weekly", type = "arithmetic")
RUS3_ret_w <- data.frame(Date = index(RUS3_ret_w), coredata(RUS3_ret_w))

Returns_weekly <- inner_join(RUS3_ret_w  , ALNY_ret_w  , by = "Date")[-1,]
colnames(Returns_weekly) <- c("Date", "ALNY", "RUS3")

start_date <- which(Returns_weekly$Date == as.Date("2019-01-04")) - 104
Returns_weekly <- Returns_weekly[-(1:(start_date)),]

# Market Return Variance

Returns_weekly$RUS3_var <- rollapply(Returns_weekly$RUS3, width = 104, FUN = var, align = "right", fill = NA)

Returns_weekly$rolling_cov <- runCov(
  x = Returns_weekly$ALNY, 
  y = Returns_weekly$RUS3, 
  n = 104
)

Returns_weekly <- Returns_weekly %>%
  mutate(
    E_Beta_weekly = rolling_cov / RUS3_var
  )

# Remove na rows
Returns_weekly <- na.omit(Returns_weekly)


ggplot(Returns_weekly, aes(x = Date, y = E_Beta_weekly)) +
  geom_line(linewidth = 1.2,color = "blue") + 
  labs(title = "Weekly Rolling Equity Beta of Alnylam Pharmaceuticals [01/2019-12/2024]", 
       x = "Date", 
       y = "Equity Beta (104 weeks window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

###  Equity Beta for monthly returns

```{r}
ALNY_ret_m <- periodReturn(Prices_xts$ALNY, period = "monthly", type = "arithmetic")
ALNY_ret_m <- data.frame(Date = index(ALNY_ret_m), coredata(ALNY_ret_m))

RUS3_ret_m <- periodReturn(Prices_xts$RUS3, period = "monthly", type = "arithmetic")
RUS3_ret_m <- data.frame(Date = index(RUS3_ret_m), coredata(RUS3_ret_m))

Returns_monthly <- inner_join(RUS3_ret_m  , ALNY_ret_m  , by = "Date")[-1,]
colnames(Returns_monthly) <- c("Date", "ALNY", "RUS3")

start_date <- which(Returns_monthly$Date == as.Date("2018-12-31")) - 60
Returns_monthly <- Returns_monthly[-(1:(start_date)),]

# Market Return Variance

Returns_monthly$RUS3_var <- rollapply(Returns_monthly$RUS3, width = 60, FUN = var, align = "right", fill = NA)

Returns_monthly$rolling_cov <- runCov(
  x = Returns_monthly$ALNY, 
  y = Returns_monthly$RUS3, 
  n = 60
)

Returns_monthly <- Returns_monthly %>%
  mutate(
    E_Beta_monthly = rolling_cov / RUS3_var
  )

# Remove na rows
Returns_monthly <- na.omit(Returns_monthly)


ggplot(Returns_monthly, aes(x = Date, y = E_Beta_monthly)) +
  geom_line(linewidth = 1.2,color = "blue") + 
  labs(title = "Monthly Rolling Equity Beta of Alnylam Pharmaceuticals [01/2019-12/2024]", 
       x = "Date", 
       y = "Equity Beta (60 months window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```






### Plot every Rolling Method 

```{r}
result1 <- Returns_daily[,c(1,6)] %>%
  full_join(Returns_weekly[,c(1,6)], by = "Date") %>%
  full_join(Returns_monthly[,c(1,6)], by = "Date")

# Interpolate missing values for each variable
result1_filled <- result1 %>%
  mutate(across(-Date, ~ zoo::na.approx(.x, na.rm = FALSE)))

# Reshape data to long format
result1_long <- result1_filled %>%
  pivot_longer(cols = -Date, names_to = "variable", values_to = "value")

# Plot using ggplot2
# Plot with separate lines for each variable
ggplot(result1_long, aes(x = Date, y = value, color = variable, group = variable)) +
  geom_line(size = 1.1, na.rm = TRUE) +  # Ignore NAs for drawing lines
  labs(title = "Rolling Equity Betas of Alnylam Pharmaceuticals [01/2019-12/2024]",
       x = "Date",
       y = "Value",
       color = "Variable") +
  theme_minimal() +
  theme(legend.position = "top")+
  scale_color_manual(values = c(E_Beta_daily = "blue",E_Beta_weekly  = "red",E_Beta_monthly  = "darkgreen"))

```

### Additional Plots

```{r}
# Rolling Plots Side by Side
plot_daily <- ggplot(Returns_daily, aes(x = Date, y = E_Beta_daily)) +
  geom_line(linewidth = 1.2, color = "blue") + 
  labs(title = "Daily Rolling Equity Beta", 
       x = "Date", 
       y = "Equity Beta (200 days window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot_weekly <- ggplot(Returns_weekly, aes(x = Date, y = E_Beta_weekly)) +
  geom_line(linewidth = 1.2, color = "red") + 
  labs(title = "Weekly Rolling Equity Beta", 
       x = "Date", 
       y = "Equity Beta (104 weeks window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot_monthly <- ggplot(Returns_monthly, aes(x = Date, y = E_Beta_monthly)) +
  geom_line(linewidth = 1.2,color = "blue") + 
  labs(title = "Monthly Rolling Equity Beta", 
       x = "Date", 
       y = "Equity Beta (60 months window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(plot_daily, plot_weekly, plot_monthly, ncol = 3)


# Table with Averages and Confidence Interval

E_Beta_d <- Returns_daily$E_Beta_daily
E_Beta_w <- Returns_weekly$E_Beta_weekly
E_Beta_m <- Returns_monthly$E_Beta_monthly

# Calculate mean and 95% confidence interval for each series (ASSUMING NORMALITY)
mean_d <- mean(E_Beta_d)
ci_d <- t.test(E_Beta_d, conf.level = 0.95)$conf.int
mean_w <- mean(E_Beta_w)
ci_w <- t.test(E_Beta_w, conf.level = 0.95)$conf.int
mean_m <- mean(E_Beta_m)
ci_m <- t.test(E_Beta_m, conf.level = 0.95)$conf.int

# Create a data frame with the results
results <- data.frame(
  Series = c("Daily", "Weekly", "Monthly"),
  Mean = round(c(mean_d, mean_w, mean_m),4),
  CI_Lower = round(c(ci_d[1], ci_w[1], ci_m[1]),4),
  CI_Upper = round(c(ci_d[2], ci_w[2], ci_m[2]),4)
)


colnames(results) <- c("Rolling Type","Avgerage Equity Beta", "CI Lower Limit", "CI Upper Limit")

results %>%
  gt() %>%
  tab_style(
    style = cell_borders(sides = "right", color = "grey80", weight = px(2)),
    locations = cells_body(columns = everything())
  ) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "grey80", weight = px(2)),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_header(title = "Average Equity Betas and Confidence Intervals for Alnylam Pharmaceuticals [01/2019-12/2024]")

```

