---
title: "Project 2 - Team 12"
output: pdf_document
date: "2025-03-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project 2 - Team 12

## Needed packages

```{r echo=TRUE, message=FALSE}
library(dplyr)
library(zoo)
library(ggplot2)
library(gridExtra)
library(frenchdata)
library(lubridate)
library(tidyr)
library(stringr)
library(zoo)
library(gt)
library(openxlsx)
library(quantmod)
library(xts)
library(TTR)
library(gridExtra)
library(tidyverse)
library(tidyquant)
library(slider)
library(broom)


```

# Exercise 6: Estimate the rolling Equity Beta 

To determine the Equity Beta, we use the Russel3000 Index, which includes Alnylam Pharmaceuticals, as market portfolio estimate. We use the following rolling formula to determine the Equity Beta (coming from the CAPM), which measures the sensitivity of a stock's returns relative to the market portfolio:


$$
\beta_t = \frac{\text{Cov}(R_{e,t}, R_{m,t})}{\text{Var}(R_{m,t})}
$$

where:

  - \( \beta_t \) is the equity beta at time t,
  - $\text{Cov}(R_e, R_m)$ is the covariance between the return on the asset $R_e$ and the return of the market portfolio $R_m$,
  - $\text{Var}(R_m)$ is the variance of the market return $R_m$.

### Download and clean the data:

```{r}
ALNY_PX <- read.xlsx("Daily_Data_Case_1.xlsx", sheet = 2) 
ALNY_PX$Date <- as.Date(ALNY_PX$Date,origin = "1899-12-30")

RUS_3_PX <- read.xlsx("Daily_Data_Case_1.xlsx", sheet = 1) 
RUS_3_PX$Date <- as.Date(RUS_3_PX$Date,origin = "1899-12-30")

# Observe NA values

ALNY_NA <- sum(is.na(ALNY_PX$Last.Price))
RUS_3_NA <- sum(is.na(RUS_3_PX$Last.Price))

# One NA in ALNY_PX --> Lin. Interpolate

na <- which(is.na(ALNY_PX$Last.Price))

ALNY_PX$Last.Price[na] <- (ALNY_PX$Last.Price[na-1] + ALNY_PX$Last.Price[na+1]) * 0.5


# Combine the data frames to match the dates

Prices <- inner_join(ALNY_PX , RUS_3_PX , by = "Date")
colnames(Prices) <- c("Date", "ALNY", "RUS3")

# Create xts for quantmod
Prices_xts <- xts(Prices[,-1], order.by = as.Date(Prices$Date))

```


### Equity Beta with daily returns

```{r}

ALNY_ret_d <- periodReturn(Prices_xts$ALNY, period = "daily", type = "arithmetic")
ALNY_ret_d <- data.frame(Date = index(ALNY_ret_d), coredata(ALNY_ret_d))

RUS3_ret_d <- periodReturn(Prices_xts$RUS3, period = "daily", type = "arithmetic")
RUS3_ret_d <- data.frame(Date = index(RUS3_ret_d), coredata(RUS3_ret_d))

Returns_daily <- inner_join(RUS3_ret_d  , ALNY_ret_d  , by = "Date")[-1,]
colnames(Returns_daily) <- c("Date", "ALNY", "RUS3")

start_date <- which(Returns_daily$Date == as.Date("2019-01-02")) - 200
Returns_daily <- Returns_daily[-(1:(start_date)),]

# Market Return Variance

Returns_daily$RUS3_var <- rollapply(Returns_daily$RUS3, width = 200, FUN = var, align = "right", fill = NA)

Returns_daily$rolling_cov <- runCov(
  x = Returns_daily$ALNY, 
  y = Returns_daily$RUS3, 
  n = 200
)

Returns_daily <- Returns_daily %>%
  mutate(
    E_Beta_daily = rolling_cov / RUS3_var
  )

# Remove na rows
Returns_daily <- na.omit(Returns_daily)

ggplot(Returns_daily, aes(x = Date, y = E_Beta_daily)) +
  geom_line(linewidth = 1.2,color = "blue") + 
  labs(title = "Daily Rolling Equity Beta of Alnylam Pharmaceuticals  [01/2019-12/2024]", 
       x = "Date", 
       y = "Equity Beta (200 days window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Equity Beta for weekly returns

```{r}

ALNY_ret_w <- periodReturn(Prices_xts$ALNY, period = "weekly", type = "arithmetic")
ALNY_ret_w <- data.frame(Date = index(ALNY_ret_w), coredata(ALNY_ret_w))

RUS3_ret_w <- periodReturn(Prices_xts$RUS3, period = "weekly", type = "arithmetic")
RUS3_ret_w <- data.frame(Date = index(RUS3_ret_w), coredata(RUS3_ret_w))

Returns_weekly <- inner_join(RUS3_ret_w  , ALNY_ret_w  , by = "Date")[-1,]
colnames(Returns_weekly) <- c("Date", "ALNY", "RUS3")

start_date <- which(Returns_weekly$Date == as.Date("2019-01-04")) - 104
Returns_weekly <- Returns_weekly[-(1:(start_date)),]

# Market Return Variance

Returns_weekly$RUS3_var <- rollapply(Returns_weekly$RUS3, width = 104, FUN = var, align = "right", fill = NA)

Returns_weekly$rolling_cov <- runCov(
  x = Returns_weekly$ALNY, 
  y = Returns_weekly$RUS3, 
  n = 104
)

Returns_weekly <- Returns_weekly %>%
  mutate(
    E_Beta_weekly = rolling_cov / RUS3_var
  )

# Remove na rows
Returns_weekly <- na.omit(Returns_weekly)


ggplot(Returns_weekly, aes(x = Date, y = E_Beta_weekly)) +
  geom_line(linewidth = 1.2,color = "blue") + 
  labs(title = "Weekly Rolling Equity Beta of Alnylam Pharmaceuticals [01/2019-12/2024]", 
       x = "Date", 
       y = "Equity Beta (104 weeks window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

###  Equity Beta for monthly returns

```{r}
ALNY_ret_m <- periodReturn(Prices_xts$ALNY, period = "monthly", type = "arithmetic")
ALNY_ret_m <- data.frame(Date = index(ALNY_ret_m), coredata(ALNY_ret_m))

RUS3_ret_m <- periodReturn(Prices_xts$RUS3, period = "monthly", type = "arithmetic")
RUS3_ret_m <- data.frame(Date = index(RUS3_ret_m), coredata(RUS3_ret_m))

Returns_monthly <- inner_join(RUS3_ret_m  , ALNY_ret_m  , by = "Date")[-1,]
colnames(Returns_monthly) <- c("Date", "ALNY", "RUS3")

start_date <- which(Returns_monthly$Date == as.Date("2018-12-31")) - 60
Returns_monthly <- Returns_monthly[-(1:(start_date)),]

# Market Return Variance

Returns_monthly$RUS3_var <- rollapply(Returns_monthly$RUS3, width = 60, FUN = var, align = "right", fill = NA)

Returns_monthly$rolling_cov <- runCov(
  x = Returns_monthly$ALNY, 
  y = Returns_monthly$RUS3, 
  n = 60
)

Returns_monthly <- Returns_monthly %>%
  mutate(
    E_Beta_monthly = rolling_cov / RUS3_var
  )

# Remove na rows
Returns_monthly <- na.omit(Returns_monthly)


ggplot(Returns_monthly, aes(x = Date, y = E_Beta_monthly)) +
  geom_line(linewidth = 1.2,color = "blue") + 
  labs(title = "Monthly Rolling Equity Beta of Alnylam Pharmaceuticals [01/2019-12/2024]", 
       x = "Date", 
       y = "Equity Beta (60 months window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```






### Plot every Rolling Method 

```{r}
result1 <- Returns_daily[,c(1,6)] %>%
  full_join(Returns_weekly[,c(1,6)], by = "Date") %>%
  full_join(Returns_monthly[,c(1,6)], by = "Date")

# Interpolate missing values for each variable
result1_filled <- result1 %>%
  mutate(across(-Date, ~ zoo::na.approx(.x, na.rm = FALSE)))

# Reshape data to long format
result1_long <- result1_filled %>%
  pivot_longer(cols = -Date, names_to = "variable", values_to = "value")

# Plot using ggplot2
# Plot with separate lines for each variable
ggplot(result1_long, aes(x = Date, y = value, color = variable, group = variable)) +
  geom_line(size = 1.1, na.rm = TRUE) +  # Ignore NAs for drawing lines
  labs(title = "Rolling Equity Betas of Alnylam Pharmaceuticals [01/2019-12/2024]",
       x = "Date",
       y = "Value",
       color = "Variable") +
  theme_minimal() +
  theme(legend.position = "top")+
  scale_color_manual(values = c(E_Beta_daily = "blue",E_Beta_weekly  = "red",E_Beta_monthly  = "darkgreen"))

```

### Additional Plots

```{r}
# Rolling Plots Side by Side
plot_daily <- ggplot(Returns_daily, aes(x = Date, y = E_Beta_daily)) +
  geom_line(linewidth = 1.2, color = "blue") + 
  labs(title = "Daily Rolling Equity Beta", 
       x = "Date", 
       y = "Equity Beta (200 days window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot_weekly <- ggplot(Returns_weekly, aes(x = Date, y = E_Beta_weekly)) +
  geom_line(linewidth = 1.2, color = "red") + 
  labs(title = "Weekly Rolling Equity Beta", 
       x = "Date", 
       y = "Equity Beta (104 weeks window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot_monthly <- ggplot(Returns_monthly, aes(x = Date, y = E_Beta_monthly)) +
  geom_line(linewidth = 1.2,color = "blue") + 
  labs(title = "Monthly Rolling Equity Beta", 
       x = "Date", 
       y = "Equity Beta (60 months window)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(plot_daily, plot_weekly, plot_monthly, ncol = 3)


# Table with Averages and Confidence Interval

E_Beta_d <- Returns_daily$E_Beta_daily
E_Beta_w <- Returns_weekly$E_Beta_weekly
E_Beta_m <- Returns_monthly$E_Beta_monthly

# Calculate mean and 95% confidence interval for each series (ASSUMING NORMALITY)
mean_d <- mean(E_Beta_d)
ci_d <- t.test(E_Beta_d, conf.level = 0.95)$conf.int
mean_w <- mean(E_Beta_w)
ci_w <- t.test(E_Beta_w, conf.level = 0.95)$conf.int
mean_m <- mean(E_Beta_m)
ci_m <- t.test(E_Beta_m, conf.level = 0.95)$conf.int

# Create a data frame with the results
results <- data.frame(
  Series = c("Daily", "Weekly", "Monthly"),
  Mean = round(c(mean_d, mean_w, mean_m),4),
  CI_Lower = round(c(ci_d[1], ci_w[1], ci_m[1]),4),
  CI_Upper = round(c(ci_d[2], ci_w[2], ci_m[2]),4)
)


colnames(results) <- c("Rolling Type","Avgerage Equity Beta", "CI Lower Limit", "CI Upper Limit")

results %>%
  gt() %>%
  tab_style(
    style = cell_borders(sides = "right", color = "grey80", weight = px(2)),
    locations = cells_body(columns = everything())
  ) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "grey80", weight = px(2)),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_header(title = "Average Equity Betas and Confidence Intervals for Alnylam Pharmaceuticals [01/2019-12/2024]")

```
## Exercise 9


### 1. Get Fama/French Weekly Data

```{r}
FF_weekly <- download_french_data("Fama/French 3 Factors [Weekly]")$subsets[, 2]$data[[1]] %>% 
  mutate(date = ymd(date)) %>% 
  filter(between(date, ymd("2017-01-01"), ymd("2024-12-31")))
```


### 2. Get ALNY Prices and Weekly Returns

```{r}
prices_alyn_weekly <- tq_get("ALNY", from = "2017-01-01", to = "2024-12-31") %>% 
  tq_transmute(select = adjusted, 
               mutate_fun = periodReturn, 
               period = "weekly", 
               type = "log", 
               col_rename = "return_weekly") %>%
  filter(between(date, ymd("2017-01-01"), ymd("2024-12-31")))

```

### 3. Join FF Factors and Stock Returns

```{r}
ff_smci <- FF_weekly %>%
  mutate(across(c(`Mkt-RF`, SMB, HML, RF), ~ .x / 100)) %>%  # convert to decimals
  left_join(prices_alyn_weekly, by = "date") %>%
  mutate(excess_return = return_weekly - RF)
```

### 4. Rolling Regression - 104 Weeks Window

```{r}
rolling_betas <- slide_dfr(
  .x = ff_smci,
  .f = ~ {
    # Only keep rows with no NAs in required columns
    window_data <- .x %>% drop_na(excess_return, `Mkt-RF`, SMB, HML)
     # Check if we have at least 104 rows after removing NAs
    if (nrow(window_data) < 104) return(tibble())
    
    fit <- lm(excess_return ~ `Mkt-RF` + SMB + HML, data = window_data)
    tibble(
      date = max(window_data$date),
      alpha = coef(fit)[1],
      beta_mkt = coef(fit)[2],
      beta_smb = coef(fit)[3],
      beta_hml = coef(fit)[4]
    )
  },
  .before = 103,
  .complete = TRUE
)
```

### 5. Plot Rolling Betas

```{r}
rolling_betas %>%
  pivot_longer(-date) %>%
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  facet_wrap(~name, scales = "free") +
  labs(x = "", y = "", title = "Rolling Fama French 3 Factors (104 Weeks)") +
  theme_bw()
```


## Exercise 10 

```{r}
rolling_comparison <- FF_weekly %>% 
  left_join(prices_alyn_weekly, by = "date") %>% 
  mutate(across(c("Mkt-RF", SMB, HML), ~ .x / 100)) %>% 
  slide(.x = ., .f = identity, .before = 103) %>% 
  keep(~ nrow(.x) == 104) %>% 
  map_dfr(~ {
    # Remove NAs before regression
    data_clean <- drop_na(.x, return_weekly, `Mkt-RF`, SMB, HML)
    
    # Skip if not enough rows
    if (nrow(data_clean) < 104) return(tibble())
    
    # Run both regressions
    ff_model <- lm(return_weekly ~ `Mkt-RF` + SMB + HML, data = data_clean)
    capm_model <- lm(return_weekly ~ `Mkt-RF`, data = data_clean)
    
    # Tidy results with confidence intervals
    ff_tidy <- tidy(ff_model, conf.int = TRUE, conf.level = 0.95) %>% mutate(model = "FF")
    capm_tidy <- tidy(capm_model, conf.int = TRUE, conf.level = 0.95) %>% mutate(model = "CAPM")
    
    # Combine results and tag with date
    bind_rows(ff_tidy, capm_tidy) %>%
      mutate(date = max(data_clean$date))
  })

```

### Plotting

```{r}
rolling_comparison %>% 
  ggplot(aes(x = date, y = estimate, color = model, fill = model)) +
  scale_color_manual(values = c("orange", "blue")) +
  scale_fill_manual(values = c("orange", "blue")) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, color = NULL),  alpha = 0.3) +
  geom_line(size = 1.1) +
  geom_hline(yintercept = 0, color = "darkred", size = 1.1, linetype = "dashed") +
  facet_wrap(~term, scales = "free_y") +
  labs(x = "", y = "", title = "Comparison of Rolling CAPM and Fama French Coefficients") +
  theme_classic() +
  theme(legend.title = element_blank())
```

## Exercise 11

### Download and clean the data:

```{r}
# Load ALNY data
ALNY_PX <- read.xlsx("Daily_Data_Case_1.xlsx", sheet = 2)
ALNY_PX$Date <- as.Date(ALNY_PX$Date, origin = "1899-12-30")

# Handle NA via linear interpolation
na <- which(is.na(ALNY_PX$Last.Price))
if (length(na) > 0) {
  ALNY_PX$Last.Price[na] <- (ALNY_PX$Last.Price[na - 1] + ALNY_PX$Last.Price[na + 1]) * 0.5
}

# Get SPY data from Yahoo Finance over the same date range
start_date <- min(ALNY_PX$Date, na.rm = TRUE)
end_date <- max(ALNY_PX$Date, na.rm = TRUE)
getSymbols("SPY", src = "yahoo", from = start_date, to = end_date)

# Convert SPY data to dataframe with adjusted prices
SPY_PX <- data.frame(Date = index(SPY), Last.Price = as.numeric(Cl(SPY))) # Ad(SPY) for the adjusted close price

# Merge ALNY and SPY prices on Date
Prices <- inner_join(ALNY_PX[, c("Date", "Last.Price")], SPY_PX, by = "Date")
colnames(Prices) <- c("Date", "ALNY", "SPY")

# Create xts object for time series analysis
Prices_xts <- xts(Prices[,-1], order.by = Prices$Date)

# Check result
head(Prices_xts)

```

## Equity Beta for weekly returns

```{r}
# Weekly returns (arithmetic)
ALNY_ret_w <- periodReturn(Prices_xts$ALNY, period = "weekly", type = "arithmetic")
ALNY_ret_w <- data.frame(Date = index(ALNY_ret_w), ALNY = coredata(ALNY_ret_w))

SPY_ret_w <- periodReturn(Prices_xts$SPY, period = "weekly", type = "arithmetic")
SPY_ret_w <- data.frame(Date = index(SPY_ret_w), SPY = coredata(SPY_ret_w))

# Join the returns
Returns_weekly <- inner_join(SPY_ret_w, ALNY_ret_w, by = "Date")[-1,]  # remove first row to align after returns
colnames(Returns_weekly) <- c("Date", "SPY", "ALNY")

# Trim to last 104 weeks from Jan 2019
start_date <- which(Returns_weekly$Date == as.Date("2019-01-04")) - 104
Returns_weekly <- Returns_weekly[-(1:(start_date)),]

# Rolling variance of market (SPY)
Returns_weekly$SPY_var <- rollapply(Returns_weekly$SPY, width = 104, FUN = var, align = "right", fill = NA)

# Rolling covariance between ALNY and SPY
Returns_weekly$rolling_cov <- runCov(
  x = Returns_weekly$ALNY, 
  y = Returns_weekly$SPY, 
  n = 104
)

# Rolling equity beta
Returns_weekly <- Returns_weekly %>%
  mutate(E_Beta_weekly = rolling_cov / SPY_var)

# Remove NA values
Returns_weekly <- na.omit(Returns_weekly)

# Plot
ggplot(Returns_weekly, aes(x = Date, y = E_Beta_weekly)) +
  geom_line(linewidth = 1.2, color = "blue") + 
  labs(
    title = "Weekly Rolling Equity Beta of Alnylam Pharmaceuticals vs SPY [2019–2024]", 
    x = "Date", 
    y = "Equity Beta (104-week window)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
