---
title: "Project 2 - Team 12"
output: pdf_document
date: "2025-03-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project 1 - Team 12

# Needed packages

```{r message=FALSE, include=FALSE}
library(dplyr)
library(zoo)
library(ggplot2)
library(gridExtra)
library(frenchdata)
library(lubridate)
library(tidyr)
library(stringr)
library(zoo)
library(gt)
library(openxlsx)
library(quantmod)
library(xts)
library(TTR)


```

# Exercise 6: Estimate the rolling Equity Beta 

To determine the Equity Beta, we use the Russel3000 Index, which includes Alnylam Pharmaceuticals, as market portfolio estimate. We use the following rolling formula to determine the Equity Beta (coming from the CAPM), which measures the sensitivity of a stock's returns relative to the market portfolio:


$$
\beta_t = \frac{\text{Cov}(R_{e,t}, R_{m,t})}{\text{Var}(R_{m,t})}
$$

where:
- \( \beta_t \) is the equity beta at time t,
- $\text{Cov}(R_e, R_m) $ is the covariance between the return on the asset $ R_e $ and the return of the market portfolio $ R_m $,
- $ \text{Var}(R_m) $ is the variance of the market return $ R_m $.

### Download and clean the data:

```{r}
ALNY_PX <- read.xlsx("Daily_Data_Case_1.xlsx", sheet = 2) 
ALNY_PX$Date <- as.Date(ALNY_PX$Date,origin = "1899-12-30")

RUS_3_PX <- read.xlsx("Daily_Data_Case_1.xlsx", sheet = 1) 
RUS_3_PX$Date <- as.Date(RUS_3_PX$Date,origin = "1899-12-30")

# Observe NA values

ALNY_NA <- sum(is.na(ALNY_PX$Last.Price))
RUS_3_NA <- sum(is.na(RUS_3_PX$Last.Price))

# One NA in ALNY_PX --> Lin. Interpolate

na <- which(is.na(ALNY_PX$Last.Price))

ALNY_PX$Last.Price[na] <- (ALNY_PX$Last.Price[na-1] + ALNY_PX$Last.Price[na+1]) * 0.5


# Combine the data frames to match the dates

Prices <- inner_join(ALNY_PX , RUS_3_PX , by = "Date")
colnames(Prices) <- c("Date", "ALNY", "RUS3")

# Create xts for quantmod
Prices_xts <- xts(Prices[,-1], order.by = as.Date(Prices$Date))

```


### Equity Beta with daily returns

```{r}

ALNY_ret_d <- periodReturn(Prices_xts$ALNY, period = "daily", type = "arithmetic")
ALNY_ret_d <- data.frame(Date = index(ALNY_ret_d), coredata(ALNY_ret_d))

RUS3_ret_d <- periodReturn(Prices_xts$RUS3, period = "daily", type = "arithmetic")
RUS3_ret_d <- data.frame(Date = index(RUS3_ret_d), coredata(RUS3_ret_d))

Returns_daily <- inner_join(RUS3_ret_d  , ALNY_ret_d  , by = "Date")[-1,]
colnames(Returns_daily) <- c("Date", "ALNY", "RUS3")

start_date <- which(Returns_daily$Date == as.Date("2019-01-02")) - 200
Returns_daily <- Returns_daily[-(1:(start_date)),]

# Market Return Variance

Returns_daily$RUS3_var <- rollapply(Returns_daily$RUS3, width = 200, FUN = var, align = "right", fill = NA)

Returns_daily$rolling_cov <- runCov(
  x = Returns_daily$ALNY, 
  y = Returns_daily$RUS3, 
  n = 200
)

Returns_daily <- Returns_daily %>%
  mutate(
    E_Beta_daily = rolling_cov / RUS3_var
  )

ggplot(Returns_daily, aes(x = Date, y = E_Beta_daily)) +
  geom_line(color = "blue") + 
  labs(title = "Daily Rolling Equity Beta of Alnylam Pharmaceuticals  [01/2019-12/2024]", 
       x = "Date", 
       y = "Equity Beta (200 days window)") +
  xlim(as.Date("2019-01-02"), as.Date("2024-12-31")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Equity Beta for weekly returns

```{r}

ALNY_ret_w <- periodReturn(Prices_xts$ALNY, period = "weekly", type = "arithmetic")
ALNY_ret_w <- data.frame(Date = index(ALNY_ret_w), coredata(ALNY_ret_w))

RUS3_ret_w <- periodReturn(Prices_xts$RUS3, period = "weekly", type = "arithmetic")
RUS3_ret_w <- data.frame(Date = index(RUS3_ret_w), coredata(RUS3_ret_w))

Returns_weekly <- inner_join(RUS3_ret_w  , ALNY_ret_w  , by = "Date")[-1,]
colnames(Returns_weekly) <- c("Date", "ALNY", "RUS3")

start_date <- which(Returns_weekly$Date == as.Date("2019-01-04")) - 104
Returns_weekly <- Returns_weekly[-(1:(start_date)),]

# Market Return Variance

Returns_weekly$RUS3_var <- rollapply(Returns_weekly$RUS3, width = 104, FUN = var, align = "right", fill = NA)

Returns_weekly$rolling_cov <- runCov(
  x = Returns_weekly$ALNY, 
  y = Returns_weekly$RUS3, 
  n = 104
)

Returns_weekly <- Returns_weekly %>%
  mutate(
    E_Beta_weekly = rolling_cov / RUS3_var
  )

ggplot(Returns_weekly, aes(x = Date, y = E_Beta_weekly)) +
  geom_line(color = "blue") + 
  labs(title = "Weekly Rolling Equity Beta of Alnylam Pharmaceuticals [01/2019-12/2024]", 
       x = "Date", 
       y = "Equity Beta (104 weeks window)") +
  xlim(as.Date("2019-01-04"), as.Date("2024-12-31")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

###  Equity Beta for monthly returns

```{r}
ALNY_ret_m <- periodReturn(Prices_xts$ALNY, period = "monthly", type = "arithmetic")
ALNY_ret_m <- data.frame(Date = index(ALNY_ret_m), coredata(ALNY_ret_m))

RUS3_ret_m <- periodReturn(Prices_xts$RUS3, period = "monthly", type = "arithmetic")
RUS3_ret_m <- data.frame(Date = index(RUS3_ret_m), coredata(RUS3_ret_m))

Returns_monthly <- inner_join(RUS3_ret_m  , ALNY_ret_m  , by = "Date")[-1,]
colnames(Returns_monthly) <- c("Date", "ALNY", "RUS3")

start_date <- which(Returns_monthly$Date == as.Date("2018-12-31")) - 60
Returns_monthly <- Returns_monthly[-(1:(start_date)),]

# Market Return Variance

Returns_monthly$RUS3_var <- rollapply(Returns_monthly$RUS3, width = 60, FUN = var, align = "right", fill = NA)

Returns_monthly$rolling_cov <- runCov(
  x = Returns_monthly$ALNY, 
  y = Returns_monthly$RUS3, 
  n = 60
)

Returns_monthly <- Returns_monthly %>%
  mutate(
    E_Beta_monthly = rolling_cov / RUS3_var
  )

ggplot(Returns_monthly, aes(x = Date, y = E_Beta_monthly)) +
  geom_line(color = "blue") + 
  labs(title = "Monthly Rolling Equity Beta of Alnylam Pharmaceuticals [01/2019-12/2024]", 
       x = "Date", 
       y = "Equity Beta (60 months window)") +
  xlim(as.Date("2018-12-31"), as.Date("2024-12-31")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

